Fast approximate Bayesian inference of HIV indicators using PCA adaptive Gauss-Hermite quadrature

### 2025-05-10 src/docs_paper-jtb/paper.Rmd
Start with revisions (a typo!)

Although the EB approach is fast, it does not take into account hyperparameter uncertainty in the latent field posterior, potentially resulting in underestimated posterior variances.
This concern could have important practical implications for use of the estimates from the Naomi model, and motivated us to look for an approach closer full Bayesian inference.
We developed a method based on adaptive Gauss-Hermite quadrature (AGHQ) extended to handle integration over many hyperparameters.
AGHQ is a quadrature method based on the theory of polynomial interpolation, and is well suited to statistical estimation problems in which the integrand is well approximated by a Gaussian multiplied by a polynomial.
For example, @bilodeau2024stochastic prove stochastic convergence rates for Bayesian posterior quantities when the normalising constant is estimated using AGHQ.
However, it is not computationally feasible to use AGHQ in greater than 20 dimensions directly, as exponentially many nodes are required.
Instead, we used principal components analysis (PCA) of the inverse curvature at the mode to find a subspace which explained most of the hyperparameter variance.
In an application to Malawi, this resulted in a grid which was tractable as it had millions of times fewer nodes than the corresponding dense grid.
Our implementation of the method makes use of the existing Naomi `TMB` template, and is immediately compatible with any model with such a template.

>>>>

Although the EB approach is fast, it does not take into account hyperparameter uncertainty in the latent field posterior, potentially resulting in underestimated posterior variances.
This concern could have important practical implications for use of the estimates from the Naomi model, and motivated us to look for an approach closer to full Bayesian inference.
We developed a method based on adaptive Gauss-Hermite quadrature (AGHQ) extended to handle integration over many hyperparameters.
AGHQ is a quadrature method based on the theory of polynomial interpolation, and is well suited to statistical estimation problems in which the integrand is well approximated by a Gaussian multiplied by a polynomial.
For example, @bilodeau2024stochastic prove stochastic convergence rates for Bayesian posterior quantities when the normalising constant is estimated using AGHQ.
However, it is not computationally feasible to use AGHQ in greater than 20 dimensions directly, as exponentially many nodes are required.
Instead, we used principal components analysis (PCA) of the inverse curvature at the mode to find a subspace which explained most of the hyperparameter variance.
In an application to Malawi, this resulted in a grid which was tractable as it had millions of times fewer nodes than the corresponding dense grid.
Our implementation of the method makes use of the existing Naomi `TMB` template, and is immediately compatible with any model with such a template.
